A Formalisation of Sturm's theorem in \isabellehol

### 2013-09-11 description/description.tex
Added user guide, updated description

\usepackage{geometry} 
\usepackage{color} 
\usepackage{graphicx} 
\usepackage{pifont} 
\usepackage[babel]{csquotes}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{pgf}
\usepackage{nicefrac}
\usepackage{enumerate}

@@

\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL}

@@

\title{A Formalisation of Sturm Chains in \isabellehol}
\author{Manuel Eberl}

@@

The goal is to formulise Sturm's theorem and its aforementioned generalisations 
in \emph{\isabellehol}. The result should be an executable function for 
determining the number of roots of a polynomial. Such a function can be 
exported to a programming language such as ML or Haskell using Isabelle's 
code generator, or it can be used to implement a decision procedure fo the 
number of roots of a polynomial directly in Isabelle.

@@

\vspace*{-5mm}
\begin{align*}
&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"
\end{align*}

@@

\section{Work outline}
We estimate that the work to be done is the following:
\begin{itemize}
\item Define, with enough generality to allow for later generalisations and optimisations, what 
      properties a general Sturm chain must fulfil
\item Prove that a chain fulfilling these requirements can indeed be used to count roots
\item Prove that the various canonical and non-canonical sturm chains fulfil these requirements
\item Prove that one can determine the signs of polynomials for sufficiently large arguments by 
      looking at the leading coefficients (this may require extension of the Polynomial library,
      since no lemmas about limits of polynomials exist so far)
\item Use these definitions and proofs to implement a decision procedure as described above and 
      integrate it with Isabelle
\end{itemize}

>>>>

\usepackage{geometry} 
\usepackage{color} 
\usepackage{graphicx} 
\usepackage{pifont} 
\usepackage[babel]{csquotes}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{pgf}
\usepackage{nicefrac}
\usepackage{enumerate}
\usepackage{xspace}

@@

\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL\xspace}

@@

\title{A Formalisation of Sturm's theorem in \isabellehol}
\author{Manuel Eberl}

@@

The goal is to formalise Sturm's theorem and its aforementioned generalisations 
in \emph{\isabellehol}. The result should be an executable function for 
determining the number of roots of a polynomial. Such a function can be 
exported to a programming language such as ML or Haskell using Isabelle's 
code generator, or it can be used to implement a decision procedure fo the 
number of roots of a polynomial directly in Isabelle.\footnote{This can only work for rational coefficients and interval bounds, of course.}

@@

\vspace*{-5mm}
\begin{align*}
&\textbf{lemma}\ "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
&\hskip15mm 1/120*x^4 + 1/24*x^4 + 1/6*x^3 - 49/16777216*x^2 - 17/2097152*x\ =\ 0\} = 3"
\end{align*}

@@

\section{Work outline}
We estimate that the work to be done is the following:
\begin{itemize}
\item Study the mathematical literature on Sturm chains
\item Define, with enough generality to allow for later generalisations and optimisations, what 
      properties a general Sturm chain must fulfil
\item Prove that a chain fulfilling these requirements can indeed be used to count roots
\item Prove that the various canonical and non-canonical sturm chains fulfil these requirements
\item Prove that one can determine the signs of polynomials for sufficiently large arguments by 
      looking at the leading coefficients (this may require extension of the Polynomial library,
      since no lemmas about limits of polynomials exist so far)
\item Use these definitions and proofs to implement a decision procedure as described above and 
      integrate it with Isabelle
\end{itemize}
